---
title: "GSE98923 co-expression analysis"
subtitle: "CEMiTool baseline - 54-sample balanced design"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    number_sections: true
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
fontsize: 11pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "92%"
)

project_dir <- Sys.getenv("CEMITOOL_PROJECT_DIR", unset = "")
if (!nzchar(project_dir)) {
  project_dir <- normalizePath("../..", winslash = "/", mustWork = TRUE)
} else {
  project_dir <- normalizePath(project_dir, winslash = "/", mustWork = TRUE)
}

samples <- read.delim(
  file.path(project_dir, "data/metadata/samples.tsv"),
  sep = "\t",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

expr <- readRDS(
  file.path(project_dir, "data/processed/expression_log2rpkm.rds")
)

cem <- readRDS(
  file.path(project_dir, "results/beta10/objects/cemitool.rds")
)

param_file <- file.path(project_dir, "results/beta10/parameters.txt")
parameters <- if (file.exists(param_file)) {
  readLines(param_file)
} else {
  character()
}

param_value <- function(key) {
  hit <- grep(paste0("^", key, "="), parameters, value = TRUE)
  if (length(hit) == 0) return(NA_character_)
  sub(paste0("^", key, "="), "", hit[[1]])
}

n_modules <- CEMiTool::nmodules(cem)
beta <- param_value("beta")
```

# Scope

This report documents the computational analysis of the balanced 54-sample subset of GEO accession **GSE98923** used in the CEMiTool baseline workflow.

The report is intended as an analysis record and potential supplementary document for a scientific manuscript. Large machine-readable matrices are not printed in full in the document; complete files are preserved separately and indexed in the supplementary manifest.

# Experimental design

The analysis contains **`r nrow(samples)` samples**, covering two cultivars, three developmental stages, three years, and three biological replicates per cultivar-stage-year combination.

```{r design-table}
design <- with(samples, table(Cultivar, Stage, Year))
knitr::kable(
  as.data.frame(design),
  caption = "Sample counts by cultivar, developmental stage, and year."
)
```

# Expression data

The processed matrix used for CEMiTool contains **`r nrow(expr)` genes** and **`r ncol(expr)` samples**.

Expression values were derived from the processed RPKM matrix distributed with GSE98923 and transformed as:

\[
\log_2(\mathrm{RPKM} + 1)
\]

The complete matrices are available under:

- `data/processed/expression_rpkm.tsv`
- `data/processed/expression_log2rpkm.tsv`

# CEMiTool analysis

CEMiTool identified **`r n_modules` modules** in the baseline analysis.

The analysis used Pearson correlation, an unsigned network, a signed TOM, a minimum module size of 30, and a fixed random seed of 1234.

Automatic beta selection did not identify a usable value in the initial baseline run. To reproduce the established 54-sample baseline, the soft-thresholding power was fixed explicitly at **beta = 10** using `set_beta = 10` and `force_beta = FALSE`. This value is specific to the present expression matrix and should be re-evaluated if the sample set, normalization, or filtering changes. The recorded beta for this run is **`r beta`**.

```{r parameters-table}
if (length(parameters) > 0) {
  p <- do.call(
    rbind,
    strsplit(parameters, "=", fixed = TRUE)
  )
  p <- as.data.frame(p, stringsAsFactors = FALSE)
  names(p) <- c("parameter", "value")
  knitr::kable(p, caption = "Recorded analysis parameters.")
}
```

# Diagnostic and analysis figures

All figures generated by the analysis are stored in `results/figures/`. PDF figures are converted to PNG copies for reliable inclusion in Word and HTML reports.

```{r figures, results='asis'}
figure_dir <- file.path(project_dir, "reports/current/assets/figures")

figure_files <- list.files(
  figure_dir,
  pattern = "\\.(png|jpg|jpeg)$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

if (length(figure_files) == 0) {
  cat("No report-compatible figure files were found.\n")
} else {
  for (f in figure_files) {
    cat("\n## ", tools::file_path_sans_ext(basename(f)), "\n\n", sep = "")
    figure_path <- file.path("..", "current", "assets", "figures", basename(f))
    figure_path <- gsub("\\\\", "/", figure_path)
    cat("![](", figure_path, "){width=92%}\n\n", sep = "")
  }
}
```

# Exported result tables

The following section provides a compact view of every tabular result exported by CEMiTool. Complete tables remain available as machine-readable files under `results/tables/`.

```{r result-tables, results='asis'}
table_files <- list.files(
  file.path(project_dir, "results/beta10/tables"),
  pattern = "\\.(tsv|csv|txt)$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

read_result_table <- function(path) {
  ext <- tolower(tools::file_ext(path))

  if (ext == "csv") {
    read.csv(
      path,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  } else {
    read.delim(
      path,
      sep = "\t",
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }
}

if (length(table_files) == 0) {
  cat("No exported result tables were found.\n")
} else {
  for (f in table_files) {
    cat("\n## ", basename(f), "\n\n", sep = "")

    x <- tryCatch(
      read_result_table(f),
      error = function(e) NULL
    )

    if (is.null(x)) {
      cat("The table could not be displayed in the report. See the original file.\n")
      next
    }

    cat(
      "Dimensions: ",
      nrow(x),
      " rows x ",
      ncol(x),
      " columns.\n\n",
      sep = ""
    )

    if (ncol(x) > 12) {
      cat(
        "The in-document preview is omitted because this matrix has ",
        ncol(x),
        " columns. The complete table remains available as a machine-readable file.\n"
      )
      next
    }

    display_n <- min(nrow(x), 40)

    print(
      knitr::kable(
        utils::head(x, display_n),
        caption = paste0(
          basename(f),
          if (nrow(x) > display_n) {
            paste0(" - first ", display_n, " rows shown")
          } else {
            ""
          }
        )
      )
    )

    if (nrow(x) > display_n) {
      cat(
        "\nThe complete table is provided as a supplementary machine-readable file.\n"
      )
    }
  }
}
```

# Supplementary data inventory

The full analysis files are indexed in `reports/supplementary/manifest.tsv`. The manifest records relative paths, file sizes, and MD5 checksums.

```{r supplement-manifest}
manifest_file <- file.path(
  project_dir,
  "reports/supplementary/manifest.tsv"
)

if (file.exists(manifest_file)) {
  manifest <- read.delim(
    manifest_file,
    sep = "\t",
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  knitr::kable(
    manifest,
    caption = "Supplementary file inventory."
  )
}
```

# Reproducibility

```{r session-info, results='asis'}
session_file <- file.path(project_dir, "results/beta10/session_info.txt")

if (file.exists(session_file)) {
  session_lines <- readLines(session_file)
  session_lines <- unlist(
    lapply(session_lines, function(line) {
      if (nchar(line) <= 88) line else strwrap(line, width = 88, exdent = 2)
    }),
    use.names = FALSE
  )
  cat("```text\n")
  cat(session_lines, sep = "\n")
  cat("\n```\n")
}
```

# Notes for manuscript use

This document is a computational analysis report, not a substitute for biological interpretation. Statements about biological mechanism, causal relationships, novelty, and comparison with previous literature should be supported separately with appropriate experimental evidence and verified references.

<!-- BEGIN POST_CEMITOOL_MODULE_STATISTICS -->

# Post-CEMiTool module statistics

This section is an additional statistical layer performed after construction of the beta-10 CEMiTool network. It is not part of the native CEMiTool report and is not stored inside the CEMiTool object.

```{r post-cemitool-load, echo=FALSE, message=FALSE, warning=FALSE}
stats_dir <- "C:/Users/jesus/OneDrive/Documentos/Cata/CEMiTool_GITHUB_SYNC/results/module_statistics_beta10"
cmp_dir <- "C:/Users/jesus/OneDrive/Documentos/Cata/CEMiTool_GITHUB_SYNC/results/comparisons"
rd <- function(x) read.delim(x, sep="\t", header=TRUE, stringsAsFactors=FALSE, check.names=FALSE)
anova <- rd(file.path(stats_dir, "module_factorial_ANOVA_typeIII.tsv"))
inter <- rd(file.path(stats_dir, "significant_cultivar_stage_interactions_FDR05.tsv"))
contr <- rd(file.path(stats_dir, "significant_cabernet_vs_pinot_stage_contrasts_FDR05.tsv"))
contr$Direction <- ifelse(contr$estimate > 0, "Cabernet Sauvignon > Pinot noir", "Cabernet Sauvignon < Pinot noir")
year <- anova[anova$Effect == "Year", , drop=FALSE]
year_sig <- year[!is.na(year$FDR) & year$FDR < 0.05, , drop=FALSE]
wide <- reshape(anova[,c("Module","Effect","FDR")], idvar="Module", timevar="Effect", direction="wide")
names(wide) <- sub("^FDR\\.", "FDR_", names(wide))
wide <- wide[order(wide$Module),,drop=FALSE]
match_file <- file.path(cmp_dir, "best_module_matches_beta7_to_beta10.tsv")
matches <- if (file.exists(match_file)) rd(match_file) else NULL
```

## Network calibration and sensitivity

Beta 10 is the primary network and beta 7 is retained as a sensitivity analysis. Module labels are arbitrary across runs, so robustness is evaluated using shared genes rather than matching labels by name.

```{r beta-robustness, echo=FALSE, results='asis'}
if (!is.null(matches)) {
  keep <- intersect(c("Module_beta7","Module_beta10","Shared_genes","Jaccard","Percent_beta7_retained"), names(matches))
  x <- matches[,keep,drop=FALSE]
  if ("Jaccard" %in% names(x)) x$Jaccard <- round(x$Jaccard,3)
  if ("Percent_beta7_retained" %in% names(x)) x$Percent_beta7_retained <- round(x$Percent_beta7_retained,1)
  print(knitr::kable(x, caption="Best beta7-to-beta10 matches based on gene membership."))
} else cat("Beta7-vs-beta10 comparison table was not available at render time.\n")
```

## Module eigengene model

For each beta-10 biological module, PC1 of standardized module-gene expression was used as the module eigengene. PC1 sign was oriented to correlate positively with mean module expression.

Model:

```text
Eigengene ~ Cultivar * Stage + Year
```

Type III ANOVA used sum-to-zero contrasts and Benjamini-Hochberg FDR.

## FDR overview by module

```{r fdr-overview, echo=FALSE, results='asis'}
for (nm in names(wide)[vapply(wide,is.numeric,logical(1))]) wide[[nm]] <- signif(wide[[nm]],4)
print(knitr::kable(wide, caption="FDR by module and model effect."))
```

## Significant Cultivar x Stage interactions

```{r interaction-table, echo=FALSE, results='asis'}
if (nrow(inter)==0) {
  cat("No module passed FDR < 0.05 for Cultivar x Stage.\n")
} else {
  s <- inter[which.min(inter$FDR),,drop=FALSE]
  cat(paste0("**",nrow(inter)," of ",length(unique(anova$Module))," modules** passed FDR < 0.05. Strongest: **",s$Module,"** (FDR=",format(s$FDR,scientific=TRUE,digits=3),").\n\n"))
  x <- inter
  x$p_value <- format(x$p_value, scientific=TRUE, digits=3)
  x$FDR <- format(x$FDR, scientific=TRUE, digits=3)
  keep <- intersect(c("Module","F value","p_value","FDR"), names(x))
  print(knitr::kable(x[,keep,drop=FALSE], caption="Significant Cultivar x Stage interactions."))
}
```

A significant interaction means that the Cabernet-versus-Pinot difference changes with developmental stage. It does not by itself demonstrate a causal relationship with skin thickness.

## Cabernet Sauvignon vs Pinot noir within stage

The contrast is Cabernet Sauvignon minus Pinot noir. Negative estimates therefore indicate a lower module eigengene in Cabernet Sauvignon.

```{r stage-contrasts, echo=FALSE, results='asis'}
if (nrow(contr)==0) {
  cat("No stage-specific contrast passed global FDR < 0.05.\n")
} else {
  x <- contr
  x$estimate <- round(x$estimate,3)
  x$p.value <- format(x$p.value, scientific=TRUE, digits=3)
  x$FDR_global <- format(x$FDR_global, scientific=TRUE, digits=3)
  keep <- intersect(c("Module","Stage","estimate","Direction","p.value","FDR_global"), names(x))
  print(knitr::kable(x[,keep,drop=FALSE], caption="Significant cultivar contrasts within stage."))
}
```

## Year effect

```{r year-effect, echo=FALSE, results='asis'}
cat(paste0("Year passed FDR < 0.05 in **",nrow(year_sig)," of ",nrow(year)," modules**.\n\n"))
if (nrow(year_sig)) {
  x <- year_sig[,c("Module","p_value","FDR"),drop=FALSE]
  x$p_value <- format(x$p_value, scientific=TRUE, digits=3)
  x$FDR <- format(x$FDR, scientific=TRUE, digits=3)
  print(knitr::kable(x, caption="Modules with significant Year effect."))
}
```

This supports keeping 2012, 2013 and 2014 explicit. The current additive model adjusts for Year but does not yet test whether the Cultivar x Stage pattern itself is stable across years.

## Interpretation boundaries and next step

- GSE98923 is berry pericarp, not isolated skin.
- There is no direct skin-thickness phenotype in this dataset.
- Only two cultivars are compared.
- Several modules have a significant Year effect.
- The baseline uses the historical GEO RPKM matrix.

The next planned statistical test is:

```text
Eigengene ~ Cultivar * Stage * Year
```

followed by Cabernet-versus-Pinot contrasts within each Stage x Year combination. Modules that remain robust can then proceed to functional annotation, hub-gene prioritization and skin-only validation.

<!-- END POST_CEMITOOL_MODULE_STATISTICS -->
