---
title: "GSE98923 co-expression analysis"
subtitle: "CEMiTool baseline - 54-sample balanced design"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    number_sections: true
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
fontsize: 11pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "92%"
)

project_dir <- normalizePath("..", winslash = "/", mustWork = TRUE)

samples <- read.delim(
  file.path(project_dir, "data/metadata/samples.tsv"),
  sep = "\t",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

expr <- readRDS(
  file.path(project_dir, "data/processed/expression_log2rpkm.rds")
)

cem <- readRDS(
  file.path(project_dir, "results/objects/cemitool.rds")
)

param_file <- file.path(project_dir, "logs/parameters.txt")
parameters <- if (file.exists(param_file)) {
  readLines(param_file)
} else {
  character()
}

param_value <- function(key) {
  hit <- grep(paste0("^", key, "="), parameters, value = TRUE)
  if (length(hit) == 0) return(NA_character_)
  sub(paste0("^", key, "="), "", hit[[1]])
}

n_modules <- CEMiTool::nmodules(cem)
beta <- param_value("beta")
```

# Scope

This report documents the computational analysis of the balanced 54-sample subset of GEO accession **GSE98923** used in the CEMiTool baseline workflow.

The report is intended as an analysis record and potential supplementary document for a scientific manuscript. Large machine-readable matrices are not printed in full in the document; complete files are preserved separately and indexed in the supplementary manifest.

# Experimental design

The analysis contains **`r nrow(samples)` samples**, covering two cultivars, three developmental stages, three years, and three biological replicates per cultivar-stage-year combination.

```{r design-table}
design <- with(samples, table(Cultivar, Stage, Year))
knitr::kable(
  as.data.frame(design),
  caption = "Sample counts by cultivar, developmental stage, and year."
)
```

# Expression data

The processed matrix used for CEMiTool contains **`r nrow(expr)` genes** and **`r ncol(expr)` samples**.

Expression values were derived from the processed RPKM matrix distributed with GSE98923 and transformed as:

\[
\log_2(\mathrm{RPKM} + 1)
\]

The complete matrices are available under:

- `data/processed/expression_rpkm.tsv`
- `data/processed/expression_log2rpkm.tsv`

# CEMiTool analysis

CEMiTool identified **`r n_modules` modules** in the baseline analysis.

The analysis used Pearson correlation, an unsigned network, a signed TOM, a minimum module size of 30, and a fixed random seed of 1234.

Automatic beta selection did not identify a usable value in the initial baseline run. To reproduce the established 54-sample baseline, the soft-thresholding power was fixed explicitly at **beta = 10** using `set_beta = 10` and `force_beta = FALSE`. This value is specific to the present expression matrix and should be re-evaluated if the sample set, normalization, or filtering changes. The recorded beta for this run is **`r beta`**.

```{r parameters-table}
if (length(parameters) > 0) {
  p <- do.call(
    rbind,
    strsplit(parameters, "=", fixed = TRUE)
  )
  p <- as.data.frame(p, stringsAsFactors = FALSE)
  names(p) <- c("parameter", "value")
  knitr::kable(p, caption = "Recorded analysis parameters.")
}
```

# Diagnostic and analysis figures

All figures generated by the analysis are stored in `results/figures/`. PDF figures are converted to PNG copies for reliable inclusion in Word and HTML reports.

```{r figures, results='asis'}
figure_dir <- file.path(project_dir, "reports/assets/figures")

figure_files <- list.files(
  figure_dir,
  pattern = "\\.(png|jpg|jpeg)$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

if (length(figure_files) == 0) {
  cat("No report-compatible figure files were found.\n")
} else {
  for (f in figure_files) {
    cat("\n## ", tools::file_path_sans_ext(basename(f)), "\n\n", sep = "")
    print(knitr::include_graphics(f))
    cat("\n")
  }
}
```

# Exported result tables

The following section provides a compact view of every tabular result exported by CEMiTool. Complete tables remain available as machine-readable files under `results/tables/`.

```{r result-tables, results='asis'}
table_files <- list.files(
  file.path(project_dir, "results/tables"),
  pattern = "\\.(tsv|csv|txt)$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

read_result_table <- function(path) {
  ext <- tolower(tools::file_ext(path))

  if (ext == "csv") {
    read.csv(
      path,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  } else {
    read.delim(
      path,
      sep = "\t",
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }
}

if (length(table_files) == 0) {
  cat("No exported result tables were found.\n")
} else {
  for (f in table_files) {
    cat("\n## ", basename(f), "\n\n", sep = "")

    x <- tryCatch(
      read_result_table(f),
      error = function(e) NULL
    )

    if (is.null(x)) {
      cat("The table could not be displayed in the report. See the original file.\n")
      next
    }

    cat(
      "Dimensions: ",
      nrow(x),
      " rows x ",
      ncol(x),
      " columns.\n\n",
      sep = ""
    )

    display_n <- min(nrow(x), 40)

    print(
      knitr::kable(
        utils::head(x, display_n),
        caption = paste0(
          basename(f),
          if (nrow(x) > display_n) {
            paste0(" - first ", display_n, " rows shown")
          } else {
            ""
          }
        )
      )
    )

    if (nrow(x) > display_n) {
      cat(
        "\nThe complete table is provided as a supplementary machine-readable file.\n"
      )
    }
  }
}
```

# Supplementary data inventory

The full analysis files are indexed in `reports/supplementary/manifest.tsv`. The manifest records relative paths, file sizes, and MD5 checksums.

```{r supplement-manifest}
manifest_file <- file.path(
  project_dir,
  "reports/supplementary/manifest.tsv"
)

if (file.exists(manifest_file)) {
  manifest <- read.delim(
    manifest_file,
    sep = "\t",
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  knitr::kable(
    manifest,
    caption = "Supplementary file inventory."
  )
}
```

# Reproducibility

```{r session-info}
session_file <- file.path(project_dir, "logs/session_info.txt")

if (file.exists(session_file)) {
  cat("```text\n")
  cat(readLines(session_file), sep = "\n")
  cat("\n```\n")
}
```

# Notes for manuscript use

This document is a computational analysis report, not a substitute for biological interpretation. Statements about biological mechanism, causal relationships, novelty, and comparison with previous literature should be supported separately with appropriate experimental evidence and verified references.
